---
title: "RepData_PA1_Shelby"
author: "Shelby Fillinger"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Peer Assessment 1 
### Load Data
```{r data}
activity <- read.csv("activity.csv")
```

## Total Steps Per Day 
```{r sum}
sum <- aggregate(x = activity$steps,   #data   
          by = list(activity$date),    #grouping          
          FUN = sum, na.rm = TRUE)    #ignore missing values
```

### Histogram of Steps Per Day 
```{r histogram}
hist(sum$x, xlab = "Steps Per Day")
```

### Mean and Median 
```{r mean}
mean <- mean(sum$x, na.rm = TRUE)   #ignore missing values
mean
median <- median(sum$x, na.rm = TRUE) #ignore missing values 
median
```

## Daily Activity Pattern 
```{r }
maxtime <- aggregate(x = activity$steps,   #data   
          by = list(activity$interval),    #grouping          
          FUN = sum, na.rm = TRUE)
```
### Time Series Daily Activity 
```{r time series}
plot(maxtime, xlab = "Interval", ylab = "Steps")
```

### Interval with Most Steps on Average 
```{r max}
t <- which.max(maxtime$x)       #row with maximum 
maxtime$Group.1[t]              #interval in relation to max row 
```

## Imputing Missing Values 
### Calculate NAs
```{r missing NAs}
sum(is.na(activity))      #sum of NAs in data set 
```
### Fill In Missing Data 
Create a new data set, impute, that will hold the data with filled NA values. The NA values are replaced with the mean number of steps taken across all days and intervals. 
```{r fill}
library(tidyr)        #import tidyr to use replace_na
imputed <- replace_na(activity, 
          list(steps = mean(activity$steps,  #replace NA in steps
                       na.rm = TRUE)))      #with the mean for steps
```

## Total Steps Per Day with Imputed Data
```{r sum2}
sum2 <- aggregate(x = imputed$steps,   #data   
          by = list(imputed$date),    #grouping          
          FUN = sum, na.rm = TRUE)    #ignore missing values
```

### Histogram of Steps Per Day with Imputed Data
```{r histogram2}
hist(sum2$x, xlab = "Steps Per Day with Imputed Data")
```

### Mean and Median with Imputed Data 
```{r mean2}
mean2 <- mean(sum2$x, na.rm = TRUE)   #ignore missing values
mean2
median2 <- median(sum2$x, na.rm = TRUE) #ignore missing values 
median2
```
Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

ANSWER

## Difference Between Weekdays and Weekends 
```{r weekday}


